Prospects of detecting gamma-ray emission from galaxy clusters: cosmic rays and dark matter annihilations

### 2011-06-15 paper.tex
minor corrections, scaling fig updated

In this paper, we will investigate in some more detail the potential
of several of the most promising galaxy clusters to produce an
annihilation gamma-ray yield which could be observable with present
and planned gamma-ray detectors. For convenience, we summarize already
in Table~\ref{tab:flux_tab} our results for the gamma-ray flux from
four promising clusters with a high gamma-ray yield. We will in the
following derive the emission from two different DM models and a
hadronic CR model to show that Virgo has the highest gamma-ray flux
from annihilating DM. If the boost from Sommerfeld enhancement (SFE)
is small, however, then the gamma-ray emission induced by CRs is
dominating most clusters with the exception of a few clusters where
Fornax had the highest flux from DM. In particular, we find Fornax and
Virgo as the prime targets for DM observations, Perseus for CR induced
emission and we additionally decide in favor of the well studied
cluster Coma for comparison. While studying these four clusters in
great detail, we complement the analysis by calculating the expected
gamma-ray fluxes of a nearby X-ray flux complete sample of galaxy
clusters from the Rosat all-sky survey (extended HIghest X-ray FLUx
Galaxy Cluster Sample, HIFLUGCS,
\citealt{2002ApJ...567..716R}).\footnote{Note that we have added the
  Virgo cluster to the sample that we nevertheless refer to as
  extended HIFLUGCS catalogue in the following.}  For previous work
related to DM in clusters, see, e.g.,
\cite{2006A&A...455...21C,2009PhRvD..80b3005J,2011arXiv1104.3530S,2011ApJ...726L...6C,2010JCAP...05..025A}.

@@

The the wave function of the lightest neutralino can be written
\begin{equation}
\tilde\chi_1=a_1\tilde B+a_2\tilde W_3+a_3 \tilde H_1^0+a_4\tilde H_2^0,
\end{equation}
with 
\begin{equation}
\sum_{i=1}^4 |a_i|^2=1.
\end{equation}
the gaugino fraction of a given neutralino is $|a_1|^2+|a_2|^2$
whereas $|a_3|^2+|a_4|^2$ is the higgsino fraction.

@@

The differential photon flux within a given solid angle $\Delta
\Omega$ along a line-of-sight (los) is given by
\begin{equation}
\label{eq:dflux}
\frac{\dd \fg}{\dd \eg} \equiv \frac{\dd^3 \ngamma}{\dd A \,\dd t\, \dd
  \egamma} = \frac{1}{2}\int_{\Delta\Omega} \dd\psi \sin\psi \frac{\dd \sg}{\dd \eg}(\psi,\eg)\,,
\end{equation}
where
\begin{equation}
\label{eq:dtildeflux}
\frac{\dd \sg}{\dd \eg}(\psi,>\eg) = \int_{\Delta\Omega} \dd\Omega \int_{\rm los}
\dd l\, q_{\rm sum}\left(\eg,r\right)\Lambda(\theta)\,,
\end{equation}
and $\sg(\psi, >\eg)$ denotes the surface brightness above the photon
energy $\eg$.  The integration along the line-of-sight $l$, in the
direction $\psi$ that the detector is pointing, is parametrized such
that the radius of the source $r=\sqrt{l^2+D_\clu^2-2 D_\clu
  l\cos\Psi}$, where $D_\clu$ is the distance from the Earth to the
center of the cluster halo and
$\cos\Psi\equiv\cos\theta\cos\psi-\cos\varphi\sin\theta\sin\psi$. The
angular integration $\dd \Omega= \sin\theta\dd \theta \,\dd \varphi$
is performed over a cone centered around $\psi$ and the opening angle
$\Delta \Omega$ is typically taken to be a few times the point spread
function (PSF) $\psf$. The limited angular resolution results in a
probability that a photon coming from a direction $\psi$' is instead
reconstructed to a direction $\psi$, where the underlying probability
distribution follow a Gaussian:
\begin{equation}
\Lambda(\theta)=\frac{1}{2\pi\psf^2}
\,\rmn{exp}\left[-\frac{\theta^2}{2\psf^2}\right]\,,
\quad \rmn{where}\quad \theta=\psi'-\psi \,.
\end{equation}
We denote the total differential source function by $q_{\rm
  sum}(\eg,r)$, where we include contributions from five main
processes; leptophilic DM annihilating \C{through
  $\chi\,\chi\to\phi\,\phi\to\{4\mu\to4e,\,4e,\,4\pi\to4e\}$ where the
  $\rmn{e}^+/\rmn{e}^-$ pairs} IC upscatter background photons
(LP-IC), leptophilic DM emitting final state radiation (LP-FSR),
supersymmetric DM BM models where annihilating neutralinos generate
$\rmn{e}^+/\rmn{e}^-$ pairs that upscatter background photons (BM-IC)
and emit a continuum as well as final state radiation (BM-Cont), and
CR proton induced $\pi^0$ that decays into gamma-rays
(CR-$\pi^0$). The source function is given by
\begin{equation}
q_\rmn{sum} (\eg,r) = \qCR(\eg,r)+
\sum_i \,q_{\rmn{sm},i}(\eg,r)\,\B_{\rmn{tot},i}(\sigv,r)
\end{equation}
where the differential CR to gamma-ray source function is denoted
by $\qCR(\eg,r)$ (see Sect.~\ref{sect:CRs} and
\cite{2010MNRAS.409..449P} for further details). The subscript $i$
runs over the gamma-ray producing DM channels and the total
differential boost factor for DM is given by:
\begin{eqnarray}
\B_{\rmn{tot},i}(r,\sigv) = \left\{\begin{array}{cc}
\B_\sfe(\sigv)\,\B_{\sub,i}(r) &\rmn{for\,\,LP}\\
\B_{\sub,i}(r) &\rmn{for\,\,BM\,.}\end{array}\right.
\end{eqnarray}
It is the product of enhancement factors from SFE $\B_\sfe(\sigv)$
(see Eq.~\ref{eq:B_sfe} and Sect.~\ref{sect:LP}) and from substructure
enhancement over the smooth halo contribution $\B_{\sub,i}(r) =
1+\rho_{\sub,i}^2(r)/\rho^2(r)$ (see
Eqs.~\ref{eq:rho_sub}-\ref{eq:xvir} and Sect.~\ref{sect:subst}).  The
DM source function from the smooth halo for each process is written in
the form:
\begin{equation}
\label{eq:q_sm}
q_{\rmn{sm},i} (\egamma,r) = \sum_j
\frac{\dd \ngammaj}{\dd E_\gamma} \Gamma_j(r)\,,
\end{equation}
where the annihilation rate density is given by 
\begin{equation}
\label{eq:ann_rate}
\Gamma_j(r) = \frac{1}{2} \left[\frac{\rho(r)}{\mx}\right]^2 
\, \sigmaannv_j\,.
\end{equation}
Here, the subscript $j$ runs over all kinematically allowed gamma-ray
producing channels, each with the spectrum $\dd
  \ngammaj /\dd\eg$ and annihilation cross section $\sigmaannv_j$.
We denote the DM mass with $\mx$ and the smooth DM density profile
with $\rho(r)$.

@@

The electrons injected from annihilating DM also suffer from diffusive
and radiative losses. Hence we have to calculate the equilibrium
spectrum of the electrons plus positrons denoted by $\dd n_\e/\dd \ee$
in Eq.~(\ref{eq:ICemiss}). We derive this stationary solution using
the cosmic ray transport equation (neglecting convection and
re-acceleration effects):
\begin{eqnarray}
\label{eq:CRdiffusion}
\frac{\partial}{\partial t}\left(\frac{\dd n_\e}{\dd \ee}\right) = 
&\nabla& \left[D\left(\ee,\bx\right)\nabla\frac{\dd n_\e}{\dd \ee}\right] + 
\frac{\partial}{\partial \ee}
\left[b\left(\ee,\bx\right) \frac{\dd n_\e}{\dd \ee}\right]
 \nonumber \\
&+& q_\e(\ee,\bx)\,,
\end{eqnarray}
where $D(\ee,\bx)$ denotes the diffusion coefficient and $b(\ee,\bx)$
the energy loss term. The source function $q_\e(\ee,\bx)$ yields the
number of electrons and positrons produced per unit time, energy and
volume element at the position $\bx$:
\begin{equation}
q_\e(\ee,r)=\sum_f\frac{\dd N_\e^f}{\dd \ee}(\ee) B_f \Gamma_f(r) \,,
\end{equation}
where the annihilation rate density $\Gamma_f(r)$ is defined in
Eq.~(\ref{eq:ann_rate}). The sum runs over the kinematically allowed
annihilation final states $f$, each with a branching ratio $B_f$ and a
differential spectrum $\dd N_\e^f/\dd \ee$ that represents the number
of electrons plus positrons resulting from an annihilation
event. \del{For neutralinos annihilating directly into electrons and
  positrons we model the spectral distribution with $\dd N_\e^f/\dd
  \ee= 2\delta(\ee-m_\chi c^2)$.} \C{We use the data from
  \cite{2011JCAP...03..019C,2011JCAP...03..051C} to compute the
spectra when neutralinos annihilate indirectly into $e^+/e^-$ pairs as
well as $\mu^+/\mu^-$ pairs in the LP model. We use \ds to compute the
$e^+/e^-$ spectra from our four BM models where only a fraction of the
annihilating neutralinos is converted into electrons and positrons
(see Sect.~\ref{sect:PF} and Fig.~\ref{fig:q_DM} for further details).}
%(see \cite{2009JCAP...01..016B} and references therein,
%also see fit).

@@

In Fig.~\ref{fig:SD_spectra} we characterize the spectral shape
through a fit to the galactic spectra presented in
\cite{2006ApJ...648L..29P}.  The figure is showing the spectra at
$r=0.03\rvir$, which is the radius there the SD energy density of a
galaxy cluster equals the energy density of the CMB black-body
distribution. Inside this central radius the SD component is
dominating, which is shown in Fig.~\ref{fig:SD_Edens} where we compare
the energy densities from different radiation fields in a galaxy
cluster with the mass $\mvir=6\times10^{14}\msun$. For this figure we
use two different profiles for the SD energy density, where the total
profile includes the contribution from the ICL, the BCG and all the
galaxies, while the galaxies are excluded in the smooth profile. To
compute the IC emission from SD, we require a non-negligible overlap
of the relativistic lepton distribution resulting from DM annihilation
and SD.  In fact, one can show with a simple order of magnitude
calculation that the overlap, $f_\rmn{IC-ol}$, of the photon field of
individual galaxies (starlight and dust emission) and the smooth DM
density is very small so that we can neglect the starlight
contribution from galaxies to the IC emission for the reminder of this
work,
\begin{eqnarray}
\label{eq:SD_overlap}
f_\rmn{IC-ol} = \frac{N_\rmn{gal} V_\rmn{gal} f_\rmn{light}}{V_\rmn{clu}}
\lesssim  \frac{N_\rmn{gal} M_\rmn{gal}}{M_\rmn{clu}\,c_\rmn{gal}^{3}}=10^{-4}.
\end{eqnarray}
Here we assume that the exponential scale height of the stellar light
is less than the scale radius of the galaxy halo, $f_\rmn{light}
\lesssim (r_\rmn{s} / r_{200,~\rmn{gal}})^3 \sim c_\rmn{gal}^{-3} \sim
10^{-3}$ and $N_\rmn{gal} M_\rmn{gal} \sim 0.1 M_\rmn{clu}$. We denote
the number of galaxies within $\rvir$ with $N_\rmn{gal}$ and the
concentration of a galaxy DM halo with $c_\rmn{gal}$.

@@

\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/fit.porter.v2.eps}
 \caption{\it Spectral dependence of radiation fields in a cluster of
   galaxies. The black dotted line of the left peak show the spectrum
   of CMB photons using a black body with a temperature of
   $2.73$~K. The crosses of the middle and right peaks represent the
   measured spectra from stars extending from the near IR to UV and
   dust at far IR wavelengths (SD), respectively, and are derived in
   \protect \cite{2006ApJ...648L..29P} for a galaxy. We normalize the
   individual SD spectrum separately using the observed luminosity
   from SD in clusters. The SD luminosity is related to the cluster
   mass through Eqs.~(\ref{eq:E_SD}-\ref{eq:N_dust}), where we use
   have used the mass $\mvir=6\times10^{14}\msun$ in this figure. We
   renormalize the SD spectra to the radius $r=0.03\rvir$, where the
   smooth energy density of the SD light (see Fig.~\ref{fig:SD_Edens})
   equals the energy density of the CMB. The red dashed lines show the
   fitted SD spectral model.}
 \label{fig:SD_spectra}
\end{figure}
\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/ucool.eps}
 \caption{\it Spatial dependence of the energy density of radiation
   fields in a cluster of galaxies. The energy density of the CMB
   (blue dotted line) is isotropic with
   $u_\rmn{cmb}=0.26\,\rmn{eV}\,\rmn{cm}^{-3}$. The energy density of
   the light from stars and dust (SD) is denoted by the red dashed
   line and the solid purple line for the total SD light and the
   smooth SD light, respectively. For comparison we show the energy
   density of the stars and a low dust model (SLD) with the light grey
   line. The SD light has been renormalized to a cluster with the mass
   $\mvir=6\times10^{14}\msun$. Finally we show the energy density of two
   magnetic field models with a central magnetic field of 3~$\mu
   G$. The magnetic field scales with the gas density to the power
   $\alpha_\rmn{B}$; green dash-dotted line ($\alpha_\rmn{B}=0.5$) and
   orange dash-dotted line ($\alpha_\rmn{B}=0.7$). Note that the SD
   radiation is dominating the energy density inside
   $\sim0.03\,\rvir$.}
 \label{fig:SD_Edens}
\end{figure}

@@

\C{We compare the calculated fluxes to gamma-ray upper limits set by
  Fermi-LAT after 18 months of observations
  \cite{2010ApJ...717L..71A}. In particular, to achieve a more
  reliable comparison we adopt the maximal spatially extended limits
  (if nothing else is stated) since the gamma-ray flux from our
  brightest clusters all have an angular extent above 1 degree on the
  sky when the boost from substructures is included. In fact, current
  Fermi flux upper limits are not derived for the kind of extended
  emission that we get with our treatment of substructures, hence the
  Fermi limits may not be adequate for some of these clusters. This and
  other recent work highlights that the improved substructure models
  imply that Fermi limits will have to be re-calculated to accommodate
  for the large source extensions. Note, however, that for most
  clusters the assumption of a point source is well justified.}

@@

In Fig.~\ref{fig:IR_comp} we show the total IC emission as well as its
individual contributions from different IC upscattered radiation
fields. The left panel shows the gamma-ray emission from the LP model
and the right panel from the $\Kp$ BM model, and for comparison we
overplot the emission expected from the CRs. Due to the flat electron
and positron spectrum resulting from the LP DM model and the smaller
mean energy of the CMB compared to the SD, we find that the
upscattered CMB photons dominate the total DM IC emission in the
energy regime below 100~GeV, while the SD dominate above this
energy. For the BM models, this transition energy is shifted towards
smaller energies since the electron and positron spectrum has a
steeper spectral distribution (see Fig.~\ref{fig:q_DM}). At the
highest gamma-ray energies of about 100~GeV and above, the IC from
starlight steepens because it probes the high energy tail of the
electrons and more importantly, it suffers from the Klein-Nishina
suppression. When substructures are present, most of their flux
resides in the outer parts of clusters. However, if we remove the
boost from substructures, the density profile of electrons and
positrons is more centrally peaked, and the relative importance of the
IC upscattered SD increases by a factor $\sim 30$ \footnote{\C{Note
    that for the leptophilic DM model this increase is much smaller
    since Sommerfeld enhancement is no longer dominated by the low
    velocity DM particles living in the subhalos.}}.  A larger
fraction of electrons (in the core) will now also cool by
Compton-upscattering SD photons which suppresses the IC-upscattered
CMB light.  For comparison, we include the contribution of IC
upscattered SD photons and to bracket the uncertainty in the SD model,
with a smaller amount of dust. In this low-dust model, we have reduced
the energy in dust by a factor 10. However, the resulting flux from
the IC upscattered dust in this model only decreases with a factor
that is slightly smaller than 10 since the IC cooling of the electrons
and positrons also decreases.

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
\includegraphics[width=0.49\columnwidth]{figures/flux.IRcomp.v13.0.1deg.1.6T.SubMass.elmu.SF700.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.IRcomp.BMv13.0.1deg.SubMass.noMW.woGal.eps}
\caption{\it Comparing the flux from different inverse Compton
  upscattered radiation fields. We show the differential inverse
  Compton emission induced by leptophilic DM in the left panel and by
  the $\Kp$ benchmark model in the right panel. The contribution from
  each individual radiation field from top line to bottom line; CMB
  (green dashed-triple-dotted), dust (purple dashed), stars (blue
  dashed-dotted). The sum of the three components are shown with the
  orange solid line.  The red solid lines show the CR induced
  gamma-ray flux. The black arrow shows the differential upper limit
  from Fermi \protect \cite{2010ApJ...717L..71A} indicating that the
  LP model assumptions such as Sommerfeld and/or substructure boost
  are in conflict with the upper limit. \del{We also show projected
    CTA point source sensitivities ($5\sigma$, 50h).}  All fluxes are
  calculated for the Fornax cluster within $\rvir$ using a point
  spread function of $0.1\deg$.\C{ For this cluster, the enhancement
    due to substructures from IC upscattered CMB and SD photons is
    730. The Sommerfeld boost is 530.}}
 \label{fig:IR_comp}
\end{minipage}
\end{figure*}

>>>>

In this paper, we will investigate in some more detail the potential
of several of the most promising galaxy clusters to produce an
annihilation gamma-ray yield which could be observable with present
and planned gamma-ray detectors. For convenience, we summarize already
in Table~\ref{tab:flux_tab} our results for the gamma-ray flux from
four promising clusters with a high gamma-ray yield. We will in the
following derive the emission from two different DM models and a
hadronic CR model to show that Virgo has the highest gamma-ray flux
from annihilating DM. If the boost from Sommerfeld enhancement (SFE)
is small, however, then the gamma-ray emission induced by CRs is
dominating most clusters with the exception of a few clusters where
Fornax had the highest flux from DM. In particular, we find Fornax and
Virgo as the prime targets for DM observations, Perseus for CR induced
emission and we additionally decide in favor of the well studied
cluster Coma for comparison. While studying these four clusters in
great detail, we complement the analysis by calculating the expected
gamma-ray fluxes of a nearby X-ray flux complete sample of galaxy
clusters from the Rosat all-sky survey (extended HIghest X-ray FLUx
Galaxy Cluster Sample, HIFLUGCS,
\citealt{2002ApJ...567..716R}).\footnote{Note that we have added the
  Virgo cluster to the sample that we nevertheless refer to as
  extended HIFLUGCS catalogue in the following.}  \C{For previous work
  related to DM in clusters, see, e.g.,
  \cite{2006A&A...455...21C,2009PhRvD..80b3005J,2011arXiv1104.3530S,2011ApJ...726L...6C,2010JCAP...05..025A}
  and for recent work on CRs in clusters see
  \cite{2010MNRAS.409..449P,2011MNRAS.410..127B,2008MNRAS.385.1211P,2009JCAP...08..002K,2010MNRAS.407.1565D}.}

@@

The wave function of the lightest neutralino can be written
\begin{equation}
\tilde\chi_1=a_1\tilde B+a_2\tilde W_3+a_3 \tilde H_1^0+a_4\tilde H_2^0\,,
\end{equation}
with 
\begin{equation}
\sum_{i=1}^4 |a_i|^2=1\,,
\end{equation}
where the gaugino fraction of a given neutralino is $|a_1|^2+|a_2|^2$
whereas $|a_3|^2+|a_4|^2$ is the higgsino fraction.

@@

The differential photon flux within a given solid angle $\Delta
\Omega$ along a line-of-sight (los) is given by
\begin{equation}
\label{eq:dflux}
\frac{\dd \fg}{\dd \eg} \equiv \frac{\dd^3 \ngamma}{\dd A \,\dd t\, \dd
  \egamma} = \frac{1}{2}\int_{\Delta\Omega} \dd\psi \sin\psi \frac{\dd \sg}{\dd \eg}(\psi,\eg)\,,
\end{equation}
where
\begin{equation}
\label{eq:dtildeflux}
\frac{\dd \sg}{\dd \eg}(\psi,>\eg) = \int_{\Delta\Omega} \dd\Omega \int_{\rm los}
\dd l\, q_{\rm sum}\left(\eg,r\right)\Lambda(\theta)\,,
\end{equation}
and $\sg(\psi, >\eg)$ denotes the surface brightness above the photon
energy $\eg$.  The integration along the line-of-sight $l$, in the
direction $\psi$ that the detector is pointing, is parametrized such
that the radius of the source $r=\sqrt{l^2+D_\clu^2-2 D_\clu
  l\cos\Psi}$, where $D_\clu$ is the distance from the Earth to the
center of the cluster halo and
$\cos\Psi\equiv\cos\theta\cos\psi-\cos\varphi\sin\theta\sin\psi$. The
angular integration $\dd \Omega= \sin\theta\dd \theta \,\dd \varphi$
is performed over a cone centered around $\psi$ and the opening angle
$\Delta \Omega$ is typically taken to be a few times the point spread
function (PSF) $\psf$. The limited angular resolution results in a
probability that a photon coming from a direction $\psi$' is instead
reconstructed to a direction $\psi$, where the underlying probability
distribution follow a Gaussian:
\begin{equation}
\Lambda(\theta)=\frac{1}{2\pi\psf^2}
\,\rmn{exp}\left[-\frac{\theta^2}{2\psf^2}\right]\,,
\quad \rmn{where}\quad \theta=\psi'-\psi \,.
\end{equation}
We denote the total differential source function by $q_{\rm
  sum}(\eg,r)$, where we include contributions from five main
processes; leptophilic DM annihilating \C{through
  $\chi\,\chi\to\phi\,\phi\to\{4e,\,4\mu\to4e,\,4\pi\to4e\}$ where the
  $\rmn{e}^+/\rmn{e}^-$ pairs} IC upscatter background photons
(LP-IC), leptophilic DM emitting final state radiation (LP-FSR),
supersymmetric DM BM models where annihilating neutralinos generate
$\rmn{e}^+/\rmn{e}^-$ pairs that upscatter background photons (BM-IC)
and emit a continuum as well as final state radiation (BM-Cont), and
CR proton induced $\pi^0$ that decays into gamma-rays
(CR-$\pi^0$). The source function is given by
\begin{equation}
q_\rmn{sum} (\eg,r) = \qCR(\eg,r)+
\sum_i \,q_{\rmn{sm},i}(\eg,r)\,\B_{\rmn{tot},i}(\sigv,r)
\end{equation}
where the differential CR to gamma-ray source function is denoted
by $\qCR(\eg,r)$ (see Sect.~\ref{sect:CRs} and
\cite{2010MNRAS.409..449P} for further details). The subscript $i$
runs over the gamma-ray producing DM channels and the total
differential boost factor for DM is given by:
\begin{eqnarray}
\B_{\rmn{tot},i}(r,\sigv) = \left\{\begin{array}{cc}
\B_\sfe(\sigv)\,\B_{\sub,i}(r) &\rmn{for\,\,LP}\\
\B_{\sub,i}(r) &\rmn{for\,\,BM\,.}\end{array}\right.
\end{eqnarray}
It is the product of enhancement factors from SFE $\B_\sfe(\sigv)$
(see Eq.~\ref{eq:B_sfe} and Sect.~\ref{sect:LP}) and from substructure
enhancement over the smooth halo contribution $\B_{\sub,i}(r) =
1+\rho_{\sub,i}^2(r)/\rho^2(r)$ \C{\footnote{Note that if the boost
    from substructures is $\gg 1$, then the Sommerfeld enhancement
    approaches the constant saturated boost of 530.}} (see
Eqs.~\ref{eq:rho_sub}-\ref{eq:xvir} and Sect.~\ref{sect:subst}).  The
DM source function from the smooth halo for each process is written in
the form:
\begin{equation}
\label{eq:q_sm}
q_{\rmn{sm},i} (\egamma,r) = \sum_j
\frac{\dd \ngammaj}{\dd E_\gamma} \Gamma_j(r)\,,
\end{equation}
where the annihilation rate density is given by 
\begin{equation}
\label{eq:ann_rate}
\Gamma_j(r) = \frac{1}{2} \left[\frac{\rho(r)}{\mx}\right]^2 
\, \sigmaannv_j\,.
\end{equation}
Here, the subscript $j$ runs over all kinematically allowed gamma-ray
producing channels, each with the spectrum $\dd
  \ngammaj /\dd\eg$ and annihilation cross section $\sigmaannv_j$.
We denote the DM mass with $\mx$ and the smooth DM density profile
with $\rho(r)$.

@@

The electrons injected from annihilating DM also suffer from diffusive
and radiative losses. Hence we have to calculate the equilibrium
spectrum of the electrons plus positrons denoted by $\dd n_\e/\dd \ee$
in Eq.~(\ref{eq:ICemiss}). We derive this stationary solution using
the cosmic ray transport equation (neglecting convection and
re-acceleration effects):
\begin{eqnarray}
\label{eq:CRdiffusion}
\frac{\partial}{\partial t}\left(\frac{\dd n_\e}{\dd \ee}\right) = 
&\nabla& \left[D\left(\ee,\bx\right)\nabla\frac{\dd n_\e}{\dd \ee}\right] + 
\frac{\partial}{\partial \ee}
\left[b\left(\ee,\bx\right) \frac{\dd n_\e}{\dd \ee}\right]
 \nonumber \\
&+& q_\e(\ee,\bx)\,,
\end{eqnarray}
where $D(\ee,\bx)$ denotes the diffusion coefficient and $b(\ee,\bx)$
the energy loss term. The source function $q_\e(\ee,\bx)$ yields the
number of electrons and positrons produced per unit time, energy and
volume element at the position $\bx$:
\begin{equation}
q_\e(\ee,r)=\sum_f\frac{\dd N_\e^f}{\dd \ee}(\ee) B_f \Gamma_f(r) \,,
\end{equation}
where the annihilation rate density $\Gamma_f(r)$ is defined in
Eq.~(\ref{eq:ann_rate}). The sum runs over the kinematically allowed
annihilation final states $f$, each with a branching ratio $B_f$ and a
differential spectrum $\dd N_\e^f/\dd \ee$ that represents the number
of electrons plus positrons resulting from an annihilation
event. \del{For neutralinos annihilating directly into electrons and
  positrons we model the spectral distribution with $\dd N_\e^f/\dd
  \ee= 2\delta(\ee-m_\chi c^2)$.} \C{We use the differential spectra
  derived from high-statistics simulations in
  \cite{2011JCAP...03..019C,2011JCAP...03..051C} to compute the
  cumulative number of electrons and positrons resulting from
  neutralinos annihilate indirectly into $e^+/e^-$ pairs as well as
  $\mu^+/\mu^-$ pairs in the LP model. We use \ds to compute the
  $e^+/e^-$ spectra from our four BM models where only a fraction of
  the annihilating neutralinos is converted into electrons and
  positrons (see Sect.~\ref{sect:PF} and Fig.~\ref{fig:q_DM} for
  further details).}
%(see \cite{2009JCAP...01..016B} and references therein,
%also see fit).

@@

In Fig.~\ref{fig:SD_spectra} we characterize the spectral shape
through a fit to the galactic spectra presented in
\cite{2006ApJ...648L..29P}.  The figure is showing the spectra at
$r=0.03\rvir$, which is the radius there the SD energy density of a
galaxy cluster equals the energy density of the CMB black-body
distribution. Inside this central radius the SD component is
dominating, which is shown in Fig.~\ref{fig:SD_Edens} where we compare
the energy densities from different radiation fields in a galaxy
cluster with the mass $\mvir=6.0\times10^{14}\msun$. For this figure we
use two different profiles for the SD energy density, where the total
profile includes the contribution from the ICL, the BCG and all the
galaxies, while the galaxies are excluded in the smooth profile. To
compute the IC emission from SD, we require a non-negligible overlap
of the relativistic lepton distribution resulting from DM annihilation
and SD.  In fact, one can show with a simple order of magnitude
calculation that the overlap, $f_\rmn{IC-ol}$, of the photon field of
individual galaxies (starlight and dust emission) and the smooth DM
density is very small so that we can neglect the starlight
contribution from galaxies to the IC emission for the reminder of this
work,
\begin{eqnarray}
\label{eq:SD_overlap}
f_\rmn{IC-ol} = \frac{N_\rmn{gal} V_\rmn{gal} f_\rmn{light}}{V_\rmn{clu}}
\lesssim  \frac{N_\rmn{gal} M_\rmn{gal}}{M_\rmn{clu}\,c_\rmn{gal}^{3}}=10^{-4}.
\end{eqnarray}
Here we assume that the exponential scale height of the stellar light
is less than the scale radius of the galaxy halo, $f_\rmn{light}
\lesssim (r_\rmn{s} / r_{200,~\rmn{gal}})^3 \sim c_\rmn{gal}^{-3} \sim
10^{-3}$ and $N_\rmn{gal} M_\rmn{gal} \sim 0.1 M_\rmn{clu}$. We denote
the number of galaxies within $\rvir$ with $N_\rmn{gal}$ and the
concentration of a galaxy DM halo with $c_\rmn{gal}$.

@@

\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/fit.porter.v2.eps}
 \caption{\it Spectral dependence of radiation fields in a cluster of
   galaxies. The black dotted line of the left peak show the spectrum
   of CMB photons using a black body with a temperature of
   $2.73$~K. The crosses of the middle and right peaks represent the
   measured spectra from stars extending from the near IR to UV and
   dust at far IR wavelengths (SD), respectively, and are derived in
   \protect \cite{2006ApJ...648L..29P} for a galaxy. We normalize the
   individual SD spectrum separately using the observed luminosity
   from SD in clusters. The SD luminosity is related to the cluster
   mass through Eqs.~(\ref{eq:E_SD}-\ref{eq:N_dust}), where we use
   have used the mass $\mvir=6.0\times10^{14}\msun$ in this figure. We
   renormalize the SD spectra to the radius $r=0.03\rvir$, where the
   smooth energy density of the SD light (see Fig.~\ref{fig:SD_Edens})
   equals the energy density of the CMB. The red dashed lines show the
   fitted SD spectral model.}
 \label{fig:SD_spectra}
\end{figure}
\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/ucool.eps}
 \caption{\it Spatial dependence of the energy density of radiation
   fields in a cluster of galaxies. The energy density of the CMB
   (blue dotted line) is isotropic with
   $u_\rmn{cmb}=0.26\,\rmn{eV}\,\rmn{cm}^{-3}$. The energy density of
   the light from stars and dust (SD) is denoted by the red dashed
   line and the solid purple line for the total SD light and the
   smooth SD light, respectively. For comparison we show the energy
   density of the stars and a low dust model (SLD) with the light grey
   line. The SD light has been renormalized to a cluster with the mass
   $\mvir=6.0\times10^{14}\msun$. Finally we show the energy density of two
   magnetic field models with a central magnetic field of 3~$\mu
   G$. The magnetic field scales with the gas density to the power
   $\alpha_\rmn{B}$; green dash-dotted line ($\alpha_\rmn{B}=0.5$) and
   orange dash-dotted line ($\alpha_\rmn{B}=0.7$). Note that the SD
   radiation is dominating the energy density inside
   $\sim0.03\,\rvir$.}
 \label{fig:SD_Edens}
\end{figure}

@@

\C{We compare the calculated fluxes to 95\% c.l. gamma-ray upper
  limits set by Fermi-LAT after 18 months of observations
  \cite{2010ApJ...717L..71A}. In particular, to achieve a more
  reliable comparison we adopt the maximal spatially extended limits
  (if nothing else is stated) since the gamma-ray flux from our
  brightest clusters all have an angular extent above 1 degree on the
  sky when the boost from substructures is included. In fact, current
  Fermi flux upper limits are not derived for the kind of extended
  emission that we get with our treatment of substructures, hence the
  Fermi limits may not be adequate for some of these clusters. This
  and other recent work highlights that the improved substructure
  models imply that Fermi limits will have to be re-calculated to
  accommodate for the large source extensions. Note, however, that for
  most clusters the assumption of a point source is well justified.}

@@

In Fig.~\ref{fig:IR_comp} we show the total IC emission as well as its
individual contributions from different IC upscattered radiation
fields. The left panel shows the gamma-ray emission from the LP model
and the right panel from the $\Kp$ BM model, and for comparison we
overplot the emission expected from the CRs. Due to the flat electron
and positron spectrum resulting from the LP DM model and the smaller
mean energy of the CMB compared to the SD, we find that the
upscattered CMB photons dominate the total DM IC emission in the
energy regime below 100~GeV, while the SD dominate above this
energy. For the BM models, this transition energy is shifted towards
smaller energies since the electron and positron spectrum has a
steeper spectral distribution (see Fig.~\ref{fig:q_DM}). At the
highest gamma-ray energies of about 100~GeV and above, the IC from
starlight steepens because it probes the high energy tail of the
electrons and more importantly, it suffers from the Klein-Nishina
suppression. When substructures are present, most of their flux
resides in the outer parts of clusters. However, if we remove the
boost from substructures, the density profile of electrons and
positrons is more centrally peaked, and the relative importance of the
IC upscattered SD increases by a factor $\sim 30$ \footnote{\C{Note
    that for the leptophilic DM model this increase is much smaller
    since Sommerfeld enhancement is no longer dominated by the low
    velocity DM particles living in the subhalos (see
    Sect.~\ref{sect:LP} for more details.}}.  A larger fraction of
electrons (in the core) will now also cool by Compton-upscattering SD
photons which suppresses the IC-upscattered CMB light.  For
comparison, we include the contribution of IC upscattered SD photons
and to bracket the uncertainty in the SD model, with a smaller amount
of dust. In this low-dust model, we have reduced the energy in dust by
a factor 10. However, the resulting flux from the IC upscattered dust
in this model only decreases with a factor that is slightly smaller
than 10 since the IC cooling of the electrons and positrons also
decreases.

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
\includegraphics[width=0.49\columnwidth]{figures/flux.IRcomp.v13.0.1deg.1.6T.SubMass.elmu.SF700.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.IRcomp.BMv13.0.1deg.SubMass.noMW.woGal.eps}
\caption{\it Comparing the flux from different inverse Compton
  upscattered radiation fields. We show the differential inverse
  Compton emission induced by leptophilic DM in the left panel and by
  the $\Kp$ benchmark model in the right panel. The contribution from
  each individual radiation field from top line to bottom line; CMB
  (green dashed-triple-dotted), dust (purple dashed), stars (blue
  dashed-dotted). The sum of the three components are shown with the
  orange solid line. The red solid lines show the CR induced gamma-ray
  flux. The black arrow shows the spatially extended differential
  upper limit from Fermi \protect \cite{2010ApJ...717L..71A}
  indicating that the LP model assumptions such as Sommerfeld and/or
  substructure boost are in conflict with the upper limit. \del{We
    also show projected CTA point source sensitivities ($5\sigma$,
    50h).}  All fluxes are calculated for the Fornax cluster within
  $\rvir$ using a point spread function of $0.1\deg$. \C{For this
    cluster, the enhancement due to substructures from IC upscattered
    CMB and SD photons is 730. The Sommerfeld boost is 530.}}
 \label{fig:IR_comp}
\end{minipage}
\end{figure*}
